Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-13 arxiv/setup.tex

Regression analysis is arguably one of the most fundamental methods in statistics and machine learning.
In many multi-task learning applications, the amount of labeled data from each dataset is usually insufficient for learning a single task.
For example, this is the case if a dataset's sample size is only a small constant factor of the feature dimension \cite{GLUE}.
In this paper, we consider multi-task learning in a \textit{high-dimensional} linear regression setting, where each dataset's sample size and the feature dimension grow linearly at a fixed ratio.
We assume that there are multiple datasets that all follow a (potentially different) linear model.
In each dataset, suppose the feature vector $x = {\Sigma}^{1/2} z$, where $z \in \real^p$ has i.i.d entries with mean zero, unit variance, and covariance $\Sigma \in\real^{p \times p}$ that is deterministic and positive semidefinite.
The high-dimensional setting is challenging but is crucial for understanding how datasets' sample sizes impact generalization performance.

@@

\begin{figure}[!t]
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.65\textwidth]{figures/mtl_model.pdf}
		\caption{A hard parameter sharing architecture}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.9\textwidth]{figures/sample_ratio_c2_400.eps}
		\caption{Varying sample size ratio}
		\label{fig_intro_sample_size_b}
	\end{subfigure}
	\caption{An illustrative example of our result:
	Consider the prediction loss of hard parameter sharing (left) for task two, given two linear regression tasks.
	Increasing task one's sample size decreases task two's prediction loss initially, but increases afterward. This phenomenon occurs due to different bias-variance tradeoffs as the sample size ratio increases. Our result provides an estimated loss (solid line) that accurately matches the empirical loss (dotted line).
	See Section \ref{sec_simulation} for the precise setting.}
	\label{fig_intro_sample_size}
\end{figure}

@@

There are two main ideas in our analysis. The proof of our first result uses a geometric intuition that hard parameter sharing finds a ``rank-$r$'' approximation of the datasets.
We carefully keep track of the concentration error between the global minimum of $f(A, B)$ and their population version.
The proof of our second result is significantly more involved because of different sample sizes and covariate shifts. Using some novel random matrix theory techniques developed recently \cite{Anisotropic}, we show that inverse of the sum of two sample covariance matrices with arbitrary population covariance matrices converge to a deterministic diagonal matrix asymptotically. Moreover, we can obtain a sharp bound on the concentration error. 
% to obtain a sharp estimate on the..., which is commonly referred to as the \emph{local law}. 
%\HZ{add several sentences on the technical insight} 
One limitation of our analysis is that for two tasks with different sample sizes, there is extra error term in the prediction loss of hard parameter sharing (cf. equation \eqref{cor_MTL_error}), which can be large for very small $n_1$. This involves a study of the asymptotic singular values distribution of a certain non-symmetric random matrix, which requires new techniques beyond the ones used in this paper. 
% plus an identity matrix,  due to lack of freeness
%\HZ{add why this is challenging}

@@

Our proof techniques use the so-called local law of random matrices \cite{erdos2017dynamical}, which is a recent development in the random matrix theory literature.
\citet{isotropic} first proved such a local law for sample covariance matrices with isotropic covariance.
This result is later extended to non-identity covariance \cite{Anisotropic} and separable covariance matrices \cite{yang2019spiked}.
%On the other hand, one may derive the asymptotic result in Theorem \ref{thm_main_RMT} with error $\oo(1)$ using the free addition of two independent random matrices in  theory .
These techniques provide the most sharp convergence rates for the asymptotic limit compared to other techniques such as free probability \cite{nica2006lectures}.
To the best of my knowledge, we are not aware of any previous result for the sum of two sample covariance matrices with arbitrary covariate shifts.

>>>>

Regression analysis is arguably one of the most fundamental methods in statistics and machine learning.
In many multi-task learning applications, the amount of labeled data from each dataset is usually insufficient for learning a single task.
For example, this is the case if a dataset's sample size is only a small constant factor of the feature dimension \cite{GLUE}.
In this paper, we consider multi-task learning in a \textit{high-dimensional} linear regression setting, where each dataset's sample size and the feature dimension grow linearly at a fixed ratio.
The high-dimensional setting is challenging but is crucial for understanding how datasets' sample sizes impact generalization performance.
We assume that there are multiple datasets that all follow a (potentially different) linear model.
In each dataset, suppose the feature vector $x = {\Sigma}^{1/2} z$, where $z \in \real^p$ has i.i.d entries with mean zero, unit variance, and covariance $\Sigma \in\real^{p \times p}$ that is deterministic and positive semidefinite.

@@

\begin{figure}[!t]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=0.5\textwidth,valign=t]{figures/mtl_model_arch.pdf}
		\caption{A hard parameter sharing architecture}
	\end{subfigure}\hfill
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=0.713\textwidth,valign=t]{figures/sample_ratio_c2_400.eps}
		\caption{Varying sample size ratio}
		\label{fig_intro_sample_size_b}
	\end{subfigure}
	\caption{An illustrative example of our result:
	Consider the prediction loss of hard parameter sharing (left) for task two, given two linear regression tasks.
	Increasing task one's sample size decreases task two's prediction loss initially, but increases afterward. This phenomenon occurs due to different bias-variance tradeoffs as the sample size ratio increases. Our result provides an estimated loss (solid line) that accurately matches the empirical loss (dotted line).
	See Section \ref{sec_simulation} for the precise setting.}
	\label{fig_intro_sample_size}
\end{figure}

@@

There are two main ideas in our analysis. The proof of our first result uses a geometric intuition that hard parameter sharing finds a ``rank-$r$'' approximation of the datasets.
We carefully keep track of the concentration error between the global minimum of $f(A, B)$ and their population version.
The proof of our second result is significantly more involved because of different sample sizes and covariate shifts. Using recently developed techniques from the random matrix theory technique \cite{Anisotropic}, we show that inverse of the sum of two sample covariance matrices with arbitrary covariate shifts converges to a deterministic diagonal matrix asymptotically (cf. Theorem \ref{thm_main_RMT}(i)). % Moreover, we can obtain a sharp bound on the concentration error.
% to obtain a sharp estimate on the..., which is commonly referred to as the \emph{local law}. 
%\HZ{add several sentences on the technical insight} 
One limitation of our analysis is that for two tasks with different sample sizes, there is extra error term in the prediction loss of hard parameter sharing (cf. equation \eqref{cor_MTL_error}), which can be large for very small $n_1$. This requires studying the spectrum of non-symmetric random matrices, which is an intriguing open question (see Section \ref{sec_conclude} for more discussion).
% plus an identity matrix,  due to lack of freeness
%\HZ{add why this is challenging}

@@

Our proof techniques use the so-called local law of random matrices \cite{erdos2017dynamical}, which is a recent development in the random matrix theory literature.
\citet{isotropic} first proved such a local law for sample covariance matrices with isotropic covariance.
\citet{Anisotropic} later extended this result to arbitrary covariances.
%On the other hand, one may derive the asymptotic result in Theorem \ref{thm_main_RMT} with error $\oo(1)$ using the free addition of two independent random matrices in  theory .
These techniques provide the most sharp convergence rates for the asymptotic limit compared to other techniques such as free probability \cite{nica2006lectures}.
To the best of our knowledge, we are not aware of any previous result for the inverse of the sum of two sample covariance matrices with arbitrary covariate shifts.
